   **Personal Movement Analysis** 
==========================

This is an Analysis Report as a part of the Project Assignment of the Reproducible Research Course. The data is collected from Personal Activity Monitoring Devices at 5-minute intervals from October and November, 2012.

## **Loading and Preprocessing the Data**

The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

The variables included are:

- **steps:** Number of steps taking in a 5-minute interval(missing values:NA)

- **date:** The date on which the measurement was taken in YYYY-MM-DD format

- **interval:** Identifier for the 5-minute interval in which measurement was taken

```{r Read&ProcessData}
data<-read.csv("activity.csv")
data$date<-as.Date(data$date)
```

## **What is mean total number of steps taken per day?**

### 1.The total number of steps taken per day

```{r totalsteps}
total.steps<-with(data,tapply(steps,date,sum,na.rm=TRUE))
df1<-data.frame(dates=names(total.steps),totalSteps=total.steps)
rownames(df1)<-seq(1,61,by=1)
head(df1)
```
### 2.Make a Histogram of total number of steps taken each day

```{r g1}
library(ggplot2)
g<-ggplot(df1,aes(x=dates,y=totalSteps))+
    geom_bar(color="darkblue",fill="lightblue",stat="identity")+
    ggtitle("Total steps taken every day",)+xlab("Date")+ylab("Total Steps")+
    theme(axis.text.x = element_text(angle=90,vjust = 0.5))
print(g)
```

### 3. The mean and median of total number of steps taken per day

```{r}
mn.day<-mean(df1$totalSteps,na.rm = TRUE)
med.day<-median(df1$totalSteps,na.rm = TRUE)
```

- Mean is : **`r mn.day`**

- Median is : **`r med.day`**

## **What is the average daily activity pattern?**

### 1. Time series plot of average number of steps for all intervals

```{r avgpattern}
avg.int.steps<-with(data,tapply(steps,interval,mean,na.rm=TRUE))
df2<-data.frame(interval=names(avg.int.steps),averageSteps=avg.int.steps)
```

```{r g2,fig.width=10}
#time-series plot
plot(df2$interval,df2$averageSteps,type="l",
     main = "Average number of steps across all intervals",xlab = "Interval",
     ylab="Average steps",lwd=2)    

```

### 2. 5-minute interval containing maximum average number of steps taken.

```{r findmax}
mx.mn.ind<-with(df2,which.max(averageSteps))
mx.mn.int<-df2[mx.mn.ind,"interval"]
mx.mn.val<-df2[mx.mn.ind,"averageSteps"]
```

Interval **`r mx.mn.int`** has maximum number of average steps **`r mx.mn.val`**..

## **Inputing missing values**

### 1.Total number of missing values(NA's)

```{r missing}
miss<-sum(is.na(data$steps))
```

Total missing values are: **`r miss`**.

### 2. Fill all the missing values in the dataset.

```{r fillmissing}
logic.na<-which(is.na(data$steps))
logic.int<-data$interval[logic.na]

missval<-sapply(logic.int,function(x,db){
    db$averageSteps[which(x==db$interval)]
},db=df2)
```

### 3. New Dataset with NA's filled with mean values of that particular interval.

```{r newdata2}
data2<-data
data2$steps[logic.na]=missval
na.val<-sum(is.na(data2$steps))
```

The total number of NA values now are: **`r na.val`**.

### 4. Histogram, mean and median of the new dataset.

```{r histnew}
total.steps.new<-with(data2,tapply(steps,date,sum,na.rm=TRUE))
df3<-data.frame(dates=names(total.steps.new),totalSteps=total.steps.new)

library(ggplot2)
g2<-ggplot(df3,aes(x=dates,y=totalSteps))+
    geom_bar(color="darkblue",fill="lightblue",stat="identity")+
    ggtitle("Total steps taken every day (NA's are replaced with mean steps on interval)",)+xlab("Date")+ylab("Total Steps")+
    theme(axis.text.x = element_text(angle=90,vjust = 0.5))
print(g2)

```

```{r}
mn.day.new<-mean(df3$totalSteps,na.rm = TRUE)
med.day.new<-median(df3$totalSteps,na.rm = TRUE)
```

- Old Mean is : **`r mn.day`**

- New Mean is : **`r mn.day.new`**

- Old Median is : **`r med.day`**

- New Median is : **`r med.day.new`**


## **Are there differences in activity patterns between weekdays and weekends?**

### 1. New Factor variable fct.week with data2.

```{r addvariable}
fct.week<-factor(ifelse(weekdays(data2$date) %in% c("Saturday","Sunday"),"weekend","weekday"))
data2<-cbind(data2,fct.week)
str(data2)
```

### 2. Panel Plot of time series interval and average steps across "weekday" and "weekend".

```{r panelplot}
spl.db<-split(data2,data2$fct.week)
reqd.values<-lapply(spl.db,function(x){
 tapply(x$steps,x$interval,mean,na.rm=TRUE)   
})

df4<-data.frame(intervals=c(names(reqd.values[[1]]),names(reqd.values[[2]])),
                mean.steps=c(reqd.values[[1]],reqd.values[[2]]),
                fct.week=(as.factor(c(rep("weekday",length(reqd.values[[1]])),
                                   rep("weekend",length(reqd.values[[2]]))))))

##plot
g3<-ggplot(data=df4,aes(x=intervals,y=mean.steps,group=fct.week))+
    facet_grid(facets = fct.week~.)+
    geom_line(color="darkblue")+
    theme(axis.text.x = element_blank())+xlab("Intervals")+
    ylab("Average Steps")+ggtitle("Mean steps across intervals for Weekdays and Weekends")
print(g3)
```